---
title: "Single Fly Descriptive Data Evaluation Sheet"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
---
<style type="text/css">
  body .main-container {
    max-width: 1800px !important;
  }
  h1.title {
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, packages, eval=FALSE, echo=FALSE}
# load packages 
library(ggplot2)
library(tidyr)
library(dygraphs)
library(grid)
library(reshape2)
library(dplyr)
library(gridExtra)
library(yaml)
library(ggsignif)
library(effsize)
library(pwr)
library(BayesFactor)
library(genefilter)  # if not available for newest version: setRepositories(addURLs = c(CRANxtras = "http://www.stats.ox.ac.uk/pub/RWin"))
library(seewave)
library(zoo)

## source the script with the functions needed for analysis
source("readXMLdatafile.R")
source("DTS_plotfunctions.R")
```

<center><h2>for time series experiment `r paste(flyname)`</h2></center>
<center>Rendered on `r format(Sys.time(), '%d %B %Y')`</center>

# Metadata

**Date and time of the experiment:** `r paste(experiment$dateTime)`  
**Raw data file (XML):** `r paste('<a href="../', basename(xml_name), '">', flyname, '</a>', sep = "")`  
**Experimenter:** `r paste(experimenter$firstname, experimenter$lastname, sep = " ")`; **ORCID:** [`r paste(experimenter$orcid)`](http://orcid.org/`r paste(experimenter$orcid)`)  
**Experiment description:** `r paste(experiment$description)`  
**Experiment duration:** `r paste(experiment$duration)`s  
**Arena type:** `r paste(experiment$arena_type)`; **Measurement device:** `r paste(experiment$meter_type)`  
**Fly description:** `r paste(fly$description)`; **FlyBase ID:** [`r paste(fly$flybase)`](http://flybase.org/reports/`r paste(fly$flybase)`)  
**Sampling rate:** `r paste(experiment$sample_rate)`Hz
```{r sr_condition, echo=FALSE}
sampleratecondition <- round(real_sample_rate)>20
```

```{r samplerate, eval=sampleratecondition, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, comment=NA, results = 'asis'}
cat("<b>Actual samplingrate: </b>", paste(signif(real_sample_rate,4)), " Hz; <b>Downsampled sampling rate (target 20 Hz):</b> ", paste(signif(down_sample_rate,4)), " Hz")
```

# Experimental Design
```{r expdesign, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 3, fig.width = (NofPeriods+1)}
    #plot table of period sequence
    periods=t(sequence)
    colnames(periods)<-sprintf("Period %s",1:NofPeriods)
    grid.table(periods)
```

# Time Traces
## Interactive Time Traces

```{r dytracesraw, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 4, fig.width = 12}    
    
    ##dyplot traces
    dytraces(traces)

```

## Interactive Time Traces (Downsampled data)

```{r dytracesdown, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 4, fig.width = 12}    
    
    ##dyplot traces
    dytraces(rawdata)

```

## Time Traces by Period (20Hz)

```{r PeriodTraces, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 10}

#add columns for PIs to sequence data
    sequence$lambda <- NA

    for(i in 1:NofPeriods){
      
      #save colors for later plotting
      if(sequence$outcome[i]==0){
        sequence$color[i]="lightyellow"
        sequence$histocolor[i]="darkgreen"
      } else {
        sequence$color[i]="orange"
        sequence$histocolor[i]="orange"
        }
     if(sequence$contingency[i]==""){
        sequence$color[i]=NA
        sequence$histocolor[i]="red"
        }
      
      ##only look at period data
      temp  <- rawdata[rawdata$period == i, ]
      keeps = c("a_pos","fly")
      period.data[[i]] <- temp[keeps] #list only arena position and fly behavior data by period
      
      if(sequence$type[i]=="fs"||sequence$type[i]=="yt"||sequence$type[i]=="color"||sequence$type[i]=="optomotorR"||sequence$type[i]=="optomotorL")
      {
        ## plot the fly behavior and arena position time traces
        fly_pos_traces(temp)
      }

      ##Calculate period histograms
      #fly behavior
      flyhistos[[i]] <- ggplot(data=temp, aes_string(temp$fly)) +
        geom_histogram(binwidth = 3, fill = sequence$histocolor[i]) +
        labs(x=paste(FlyBehavior, "[arb units]"), y="frequency") +
        xlim(maxfly) +
        ggtitle(paste(flyname, "Period", i))
      
      #arena position
      if(sequence$type[i]=="fs"||sequence$type[i]=="color")
      {
        poshistos[[i]] <- ggplot(data=temp, aes_string(temp$a_pos)) +
          geom_histogram(binwidth=10, fill = sequence$histocolor[i]) +
          labs(x="Arena Position [degrees*10]", y="frequency") +
          xlim(-1800,1800) +
          ggtitle(paste(flyname, "Period", i))
      }
      
      
      ##calculate PIs
      ##create data.frame for adding PIs if it doesn't eist, yet
      if(!exists("PIprofile")){PIprofile <- data.frame(matrix(ncol = NofPeriods))}
      
      if(sequence$type[i]=="fs"||sequence$type[i]=="color") #for arena position
      {
        t1 = sum(abs(temp$a_pos) >= 450 & abs(temp$a_pos) <= 1350)
        t2 = nrow(temp)-t1
        sequence$lambda[i] = (t1-t2)/(t1+t2)
        if (sequence$contingency[i] == '2_4_Q'){sequence$lambda[i]=-sequence$lambda[i]}
        if(l==1){PIprofile[1,i]=sequence$lambda[i]}
      } else if (sequence$type[i]=="yt" || sequence$type[i]=="sw_blue" || sequence$type[i]=="sw_green") #calculate torque PIs
      {
        t2 = table(sign(temp$fly))[1]
        t1 = nrow(temp)-t2
        sequence$lambda[i] = (t1-t2)/(t1+t2)
        if (sequence$contingency[i] == 'right_torque'){sequence$lambda[i]=-sequence$lambda[i]}
        if(l==1){PIprofile[1,i]=sequence$lambda[i]}
      }
    } #for Number of Periods
    
```

# Histograms
## Fly Behavior Histograms

```{r FlyHistogram, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8}    
    
    ## plot position and fly histograms ##
    # add fly behavior pooled histogram
    flyhistos[[NofPeriods+1]] <- ggplot(data=rawdata, aes_string(rawdata$fly)) + 
      geom_histogram(binwidth=3) + 
      labs(x=paste(FlyBehavior, "[arb units]"), y="frequency") + 
      xlim(-maxfly) +
      ggtitle(paste(flyname, "total"))
    
    # add pooled position histogram - if there are any periods with meaningful position data
    if ('fs' %in% sequence$type || 'color' %in% sequence$type) {
      poshistos[[NofPeriods+1]] <- ggplot(data=rawdata, aes_string(rawdata$a_pos)) + 
      geom_histogram(binwidth=10) +
      labs(x="position [degrees*10]", y="frequency") + 
      xlim(-1800,1800) +
      ggtitle(paste(flyname, "total"))
    }
    
    ##write histograms
    #fly behavior
    multiplot(plotlist = flyhistos, cols=round(NofPeriods/5))
```

```{r PositionHistogram, eval=('fs' %in% sequence$type || 'color' %in% sequence$type), echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    
      cat("\n## Arena Position Histograms\n")
      multiplot(plotlist = Filter(Negate(is.null), poshistos), cols=round(sum(poshistos!="NULL")/5)) #only print position histograms if there are any

```

<!-- Evaluate optomotor data if they are present -->

```{r om, echo=FALSE}
om <- any(grepl("optomotor", sequence$type)==TRUE)    ###determine if there are optomotor periods
if (om & any(!grepl("optomotor", sequence$type)==TRUE)){   ###if there are non-optomotor periods...
  if (grepl("optomotor", sequence$type[1]) & grepl("optomotor", tail(sequence$type, 1))){ ###...and the opto periods are in the beginning and the end
    beforeafter = TRUE
    om = FALSE
  }
}
```

`r if (om | beforeafter) '# Optomotor evaluations'`
`r if (om | beforeafter) '## Optomotor traces (right/left)'`

```{r OptomotorTraces, eval=om, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

### extract optomotor traces and generate dataframe for right (clockwise) turning arena
rightOMperiods=which(grepl("optomotorR", sequence$type)) #collect the right turning OM periods
rightOMdata<-filter(rawdata, rawdata$period %in% rightOMperiods) #extract only right turning arena periods
rightOMdata <- rightOMdata %>% select(-c(a_pos,date)) #drop unnecessary columns
rightOMdata$time=ave(rightOMdata$period, rightOMdata$period, FUN=seq_along) #match the time values to start at each period start
rightOMdata$time=(rightOMdata$time-1)*50 #make 20Hz data into ms time scale
rightOMdata$period=as.factor(rightOMdata$period)
ggplot(data = rightOMdata, aes(x=time/1000, y=fly)) + 
  geom_hline(yintercept = 0, color="black") +
  geom_line(aes(group=period, colour=period), size=1) + 
  geom_smooth(method="loess", span = 0.1) +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") + 
  ggtitle(paste("Right (clockwise) arena rotations -", flyname)) +
  theme_classic()+
  scale_x_continuous(expand = expand_scale(add = 0))+
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))

### extract optomotor traces and generate dataframe for left turning (counter-clockwise) arena
leftOMperiods=which(grepl("optomotorL", sequence$type)) #collect the left turning OM periods
leftOMdata<-filter(rawdata, rawdata$period %in% leftOMperiods)  #extract only left turning arena periods
leftOMdata <- leftOMdata %>% select(-c(a_pos,date))  #drop unnecessary columns
leftOMdata$time=ave(leftOMdata$period, leftOMdata$period, FUN=seq_along) #match the time values to start at each period start
leftOMdata$time=((leftOMdata$time)*50)+max(rightOMdata$time) #make 20Hz data into ms time scale, starting after right turning data
leftOMdata$period=as.factor(leftOMdata$period)
ggplot(data = leftOMdata, aes(x=time/1000, y=fly)) + 
  geom_hline(yintercept = 0, color="black") +
  geom_line(aes(group=period, colour=period), size=1) + 
  geom_smooth(method="loess", span = 0.1) +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") +
  ggtitle(paste("Left (counter-clockwise) arena rotations -", flyname)) +
  theme_classic() +
  scale_x_continuous(expand = expand_scale(add = 0))+
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))

### generate averaged optomotor responses over both rotation directions
tempOMdata=rbind(rightOMdata,leftOMdata) #combine left and right OMdata in one dataframe
tempOMdata=aggregate(tempOMdata$fly,FUN=mean,by=list(tempOMdata$time)) #average across periods
colnames(tempOMdata) <- c("time", as.character(flyname)) #set dataframe names

if(!exists("OMdata"))
{
  OMdata=tempOMdata
} else {
  OMdata[[as.character(flyname)]]=tempOMdata[[as.character(flyname)]]
}
```

`r if (beforeafter) '### Optomotor traces at start'`

```{r OptomotorTracesBefore, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

### extract optomotor traces and generate dataframe for right (clockwise) turning arena
rightOMbefore=which(grepl("optomotorR", head(sequence$type))) #collect the right turning OM periods in the first half of the experiment
rightOMdataBefore<-filter(rawdata, rawdata$period %in% rightOMperiods) #extract only right turning arena periods
rightOMdataBefore <- rightOMdataBefore %>% select(-c(a_pos,date)) #drop unnecessary columns
rightOMdataBefore$time=ave(rightOMdataBefore$period, rightOMdataBefore$period, FUN=seq_along) #match the time values to start at each period start
rightOMdataBefore$time=(rightOMdataBefore$time-1)*50 #make 20Hz data into ms time scale
rightOMdataBefore$period=as.factor(rightOMdataBefore$period)
ggplot(data = rightOMdataBefore, aes(x=time/1000, y=fly)) + 
  geom_hline(yintercept = 0, color="black") +
  geom_line(aes(group=period, colour=period), size=1) + 
  geom_smooth(method="loess", span = 0.1) +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") + 
  ggtitle(paste("Right (clockwise) arena rotations -", flyname)) +
  theme_classic()+
  scale_x_continuous(expand = expand_scale(add = 0))+
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))

### extract optomotor traces and generate dataframe for left turning (counter-clockwise) arena
leftOMbefore=which(grepl("optomotorL", head(sequence$type))) #collect the left turning OM periods in the first half of the experiment
leftOMdataBefore<-filter(rawdata, rawdata$period %in% leftOMperiods)  #extract only left turning arena periods
leftOMdataBefore <- leftOMdataBefore %>% select(-c(a_pos,date))  #drop unnecessary columns
leftOMdataBefore$time=ave(leftOMdataBefore$period, leftOMdataBefore$period, FUN=seq_along) #match the time values to start at each period start
leftOMdataBefore$time=((leftOMdataBefore$time)*50)+max(rightOMdataBefore$time) #make 20Hz data into ms time scale, starting after right turning data
leftOMdataBefore$period=as.factor(leftOMdataBefore$period)
ggplot(data = leftOMdataBefore, aes(x=time/1000, y=fly)) + 
  geom_hline(yintercept = 0, color="black") +
  geom_line(aes(group=period, colour=period), size=1) + 
  geom_smooth(method="loess", span = 0.1) +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") +
  ggtitle(paste("Left (counter-clockwise) arena rotations -", flyname)) +
  theme_classic() +
  scale_x_continuous(expand = expand_scale(add = 0))+
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))

### generate averaged optomotor responses over both rotation directions
tempOMdataBefore=rbind(rightOMdataBefore,leftOMdataBefore) #combine left and right OMdata in one dataframe
tempOMdataBefore=aggregate(tempOMdataBefore$fly,FUN=mean,by=list(tempOMdataBefore$time)) #average across periods
colnames(tempOMdataBefore) <- c("time", as.character(flyname)) #set dataframe names

if(!exists("OMdataBefore"))
{
  OMdataBefore=tempOMdataBefore
} else {
  OMdataBefore[[as.character(flyname)]]=tempOMdataBefore[[as.character(flyname)]]
}
```

`r if (beforeafter) '### Optomotor traces at end'`

```{r OptomotorTracesAfter, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

### extract optomotor traces and generate dataframe for right (clockwise) turning arena
rightOMafter=which(grepl("optomotorR", tail(sequence$type))) #collect the right turning OM periods in the last half of the experiment
rightOMdataAfter<-filter(rawdata, rawdata$period %in% rightOMperiods) #extract only right turning arena periods
rightOMdataAfter <- rightOMdataAfter %>% select(-c(a_pos,date)) #drop unnecessary columns
rightOMdataAfter$time=ave(rightOMdataAfter$period, rightOMdataAfter$period, FUN=seq_along) #match the time values to start at each period start
rightOMdataAfter$time=(rightOMdataAfter$time-1)*50 #make 20Hz data into ms time scale
rightOMdataAfter$period=as.factor(rightOMdataAfter$period)
ggplot(data = rightOMdataAfter, aes(x=time/1000, y=fly)) + 
  geom_hline(yintercept = 0, color="black") +
  geom_line(aes(group=period, colour=period), size=1) + 
  geom_smooth(method="loess", span = 0.1) +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") + 
  ggtitle(paste("Right (clockwise) arena rotations -", flyname)) +
  theme_classic()+
  scale_x_continuous(expand = expand_scale(add = 0))+
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))

### extract optomotor traces and generate dataframe for left turning (counter-clockwise) arena
leftOMafter=which(grepl("optomotorL", tail(sequence$type))) #collect the left turning OM periods in the last half of the experiment
leftOMdataAfter<-filter(rawdata, rawdata$period %in% leftOMperiods)  #extract only left turning arena periods
leftOMdataAfter <- leftOMdataAfter %>% select(-c(a_pos,date))  #drop unnecessary columns
leftOMdataAfter$time=ave(leftOMdataAfter$period, leftOMdataAfter$period, FUN=seq_along) #match the time values to start at each period start
leftOMdataAfter$time=((leftOMdataBefore$time)*50)+max(rightOMdataAfter$time) #make 20Hz data into ms time scale, starting after right turning data
leftOMdataAfter$period=as.factor(leftOMdataAfter$period)
ggplot(data = leftOMdataAfter, aes(x=time/1000, y=fly)) + 
  geom_hline(yintercept = 0, color="black") +
  geom_line(aes(group=period, colour=period), size=1) + 
  geom_smooth(method="loess", span = 0.1) +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") +
  ggtitle(paste("Left (counter-clockwise) arena rotations -", flyname)) +
  theme_classic() +
  scale_x_continuous(expand = expand_scale(add = 0))+
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))

### generate averaged optomotor responses over both rotation directions
tempOMdataAfter=rbind(rightOMdataAfter,leftOMdataAfter) #combine left and right OMdata in one dataframe
tempOMdataAfter=aggregate(tempOMdataAfter$fly,FUN=mean,by=list(tempOMdataAfter$time)) #average across periods
colnames(tempOMdataAfter) <- c("time", as.character(flyname)) #set dataframe names

if(!exists("OMdataAfter"))
{
  OMdataAfter=tempOMdataAfter
} else {
  OMdataAfter[[as.character(flyname)]]=tempOMdataAfter[[as.character(flyname)]]
}
```

`r if (om | beforeafter) '## Averaged optomotor response'`

```{r OptomotorAverage, eval=om, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

ggplot(data=OMdata, aes(x=OMdata$time/1000, y=OMdata[[as.character(flyname)]])) + 
  geom_rect(aes(xmin = mean(OMdata$time/1000),xmax = Inf,ymin = -Inf, ymax = Inf),fill=("grey"), alpha = 0.01)+
  geom_hline(yintercept = 0) +
  geom_line() +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") +
  theme_classic() + 
  annotate("text", -Inf, Inf, label = "Right (clockwise) arena rotations", hjust = -0.1, vjust = 1.3)+
  annotate("text", Inf, -Inf, label = "Left (counter-clockwise) arena rotations", hjust =1.1, vjust = -1.5)+ 
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))+
  scale_x_continuous(expand = expand_scale(add = 0))
```

`r if (beforeafter) '### Averaged optomotor response at start'`

```{r OptomotorAverageBefore, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

ggplot(data=OMdataBefore, aes(x=OMdataBefore$time/1000, y=OMdataBefore[[as.character(flyname)]])) + 
  geom_rect(aes(xmin = mean(OMdataBefore$time/1000),xmax = Inf,ymin = -Inf, ymax = Inf),fill=("grey"), alpha = 0.01)+
  geom_hline(yintercept = 0) +
  geom_line() +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") +
  theme_classic() + 
  annotate("text", -Inf, Inf, label = "Right (clockwise) arena rotations", hjust = -0.1, vjust = 1.3)+
  annotate("text", Inf, -Inf, label = "Left (counter-clockwise) arena rotations", hjust =1.1, vjust = -1.5)+ 
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))+
  scale_x_continuous(expand = expand_scale(add = 0))
```

`r if (beforeafter) '### Averaged optomotor response at end'`

```{r OptomotorAverageAfter, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

ggplot(data=OMdataAfter, aes(x=OMdataAfter$time/1000, y=OMdataAfter[[as.character(flyname)]])) + 
  geom_rect(aes(xmin = mean(OMdataAfter$time/1000),xmax = Inf,ymin = -Inf, ymax = Inf),fill=("grey"), alpha = 0.01)+
  geom_hline(yintercept = 0) +
  geom_line() +
  ylab("Optomotor Resoponse [rel. units]")+ 
  xlab("Time [s]") +
  theme_classic() + 
  annotate("text", -Inf, Inf, label = "Right (clockwise) arena rotations", hjust = -0.1, vjust = 1.3)+
  annotate("text", Inf, -Inf, label = "Left (counter-clockwise) arena rotations", hjust =1.1, vjust = -1.5)+ 
  theme(panel.grid.major.x = element_blank(),panel.grid.major.y = element_line( size=.1, color="grey"))+
  scale_x_continuous(expand = expand_scale(add = 0))
```

`r if (om | beforeafter) '## Further optomotor evaluations'`

```{r moreOptomotor, eval=om, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

###generate traces to fit nonlinear model for parameter estimation
# Find the midpoint in left and right turning in averaged traces (OMdata)
OMmidpoint = OMdata$time[nrow(OMdata)]/2

#right turning traces
x1=OMdata$time[OMdata$time<=OMmidpoint]                     #time
y=OMdata[[as.character(flyname)]][OMdata$time<=OMmidpoint]  #mean right OM response
y=rollapply(y, 20,mean, fill="extend")                      #apply rolling average of 20 data points (1s)
xy <-sortedXyData(x1,y)                                     #create sortedXYdata object

LinMod = lm(y ~ x1, data = xy)                              #fit linear model to OMdata right turning

#plot(xy)
#abline(LinMod)
#LinMod$coefficients[2]

if(abs(LinMod$coefficients[2])<0.0001){                     #if the linear model shows hardly any slope, set asymptote as mean of all values
  right=c(b0=0, b1=mean(y), lrc=-1000)
  } else {
right<-tryCatch(                                            #extract the right turning coefficients into a vector. Set vector to zero for errors
    expr = {
                    right<-NLSstAsymptotic(xy)
                     },
    error = function(e){
                      right=c(b0=0, b1=0, lrc=-1000)        #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(right)
                     },
    warning = function(w){
                      right=c(b0=0, b1=0, lrc=-1000)        #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(right)
                     },
    finally = {}
)}

#left turning traces
x1=OMdata$time[OMdata$time>OMmidpoint]                      #time
x1=x1-x1[1]                                                 #set time to start from zero
y=OMdata[[as.character(flyname)]][OMdata$time>OMmidpoint]   #mean left OM response
y=rollapply(y, 20,mean, fill="extend")                      #apply rolling average of 20 data points (1s)
xy <-sortedXyData(x1,y)                                     #create sortedXYdata object

LinMod = lm(y ~ x1, data = xy)                              #fit linear model to OMdata left turning

#plot(xy)
#abline(LinMod)
#LinMod$coefficients[2]

if(abs(LinMod$coefficients[2])<0.0001){                     #if the linear model shows hardly any slope, set asymptote as mean of all values
  left=c(b0=0, b1=mean(y), lrc=-1000)
  } else {
left<-tryCatch(                                             #extract the left turning coefficients into a vector. Set vector to zero for errors
    expr = {
                    left<-NLSstAsymptotic(xy)
                     },
    error = function(e){
                      left=c(b0=0, b1=0, lrc=-1000)         #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(left)
                     },
    warning = function(w){
                      left=c(b0=0, b1=0, lrc=-1000)         #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(left)
                     },
    finally = {}
)}


###calculate averages and asymmetries to three significant digits and put them into a vector
rightasymp = right[1]+right[2]
leftasymp = left[1]+left[2]
#sanity checks
if(leftasymp < maxfly[1]){leftasymp=0}
if(rightasymp > maxfly[2]){rightasymp=0}

if (leftasymp==0 & rightasymp==0){aveOMcoeff=c(0,0,0)
} else {
      aveOMcoeff=c(signif((abs(rightasymp)+abs(leftasymp))/2, 3),                                #mean asymptote
                   signif(((exp(right[3])+exp(left[3]))/2)*1000, 3),                             #mean rate constant [s]
                   signif((abs(rightasymp)-abs(leftasymp))/(abs(rightasymp)+abs(leftasymp)), 3)) #asymmetry index
}

###make vector into a dataframe for easy exporting and plotting
tempOMparams <- as.data.frame(rbind(as.numeric(aveOMcoeff))) #convert to dataframe
names(tempOMparams) = c("asymptote", "rate constant","AI")        #set column names
rownames(tempOMparams)[1]=as.character(flyname)              #set row names to flynames
kable(tempOMparams)                                          #plot table with data

###collect all coefficients in dataframe for the group
if(!exists("OMparams"))
{
  OMparams=tempOMparams
} else {
  OMparams[as.character(flyname),]=tempOMparams
}

```

`r if (beforeafter) '### Optomotor parameters at start'`

```{r OMparamsBefore, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

###generate traces to fit nonlinear model for parameter estimation
# Find the midpoint in left and right turning in averaged traces (OMdata)
OMmidBefore = OMdataBefore$time[nrow(OMdataBefore)]/2

#right turning traces
x1=OMdataBefore$time[OMdataBefore$time<=OMmidBefore]                     #time
y=OMdataBefore[[as.character(flyname)]][OMdataBefore$time<=OMmidBefore]  #mean right OM response
y=rollapply(y, 40,mean, fill="extend")                                   #apply rolling average of 20 data points (2s)
xy <-sortedXyData(x1,y)                                                  #create sortedXYdata object

LinMod = lm(y ~ x1, data = xy)                                           #fit linear model to OMdata right turning

#plot(xy)
#abline(LinMod)
#LinMod$coefficients[2]

if(abs(LinMod$coefficients[2])<0.0001){                     #if the linear model shows hardly any slope, set asymptote as mean of all values
  right=c(b0=0, b1=mean(y), lrc=-1000)
  } else {
right<-tryCatch(                                            #extract the right turning coefficients into a vector. Set vector to zero for errors
    expr = {
                    right<-NLSstAsymptotic(xy)
                     },
    error = function(e){
                      right=c(b0=0, b1=0, lrc=-1000)        #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(right)
                     },
    warning = function(w){
                      right=c(b0=0, b1=0, lrc=-1000)        #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(right)
                     },
    finally = {}
)}

#left turning traces
x1=OMdataBefore$time[OMdataBefore$time>OMmidBefore]                      #time
x1=x1-x1[1]                                                              #set time to start from zero
y=OMdataBefore[[as.character(flyname)]][OMdataBefore$time>OMmidBefore]   #mean left OM response
y=rollapply(y, 40,mean, fill="extend")                                   #apply rolling average of 20 data points (2s)
xy <-sortedXyData(x1,y)                                                  #create sortedXYdata object

LinMod = lm(y ~ x1, data = xy)                              #fit linear model to OMdata left turning

#plot(xy)
#abline(LinMod)
#LinMod$coefficients[2]

if(abs(LinMod$coefficients[2])<0.0001){                     #if the linear model shows hardly any slope, set asymptote as mean of all values
  left=c(b0=0, b1=mean(y), lrc=-1000)
  } else {
left<-tryCatch(                                             #extract the left turning coefficients into a vector. Set vector to zero for errors
    expr = {
                    left<-NLSstAsymptotic(xy)
                     },
    error = function(e){
                      left=c(b0=0, b1=0, lrc=-1000)         #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(left)
                     },
    warning = function(w){
                      left=c(b0=0, b1=0, lrc=-1000)         #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(left)
                     },
    finally = {}
)}


###calculate averages and asymmetries to three significant digits and put them into a vector
rightasymp = right[1]+right[2]
leftasymp = left[1]+left[2]
#sanity checks
if(leftasymp < maxfly[1]){leftasymp=0}
if(rightasymp > maxfly[2]){rightasymp=0}

if (leftasymp==0 & rightasymp==0){aveOMcoeff=c(0,0,0)
} else {
      aveOMcoeffBefore=c(signif((abs(rightasymp)+abs(leftasymp))/2, 3),                                #mean asymptote
                   signif(((exp(right[3])+exp(left[3]))/2)*1000, 3),                             #mean rate constant [s]
                   signif((abs(rightasymp)-abs(leftasymp))/(abs(rightasymp)+abs(leftasymp)), 3)) #asymmetry index
}

###make vector into a dataframe for easy exporting and plotting
tempOMparamsBefore <- as.data.frame(rbind(as.numeric(aveOMcoeffBefore))) #convert to dataframe
names(tempOMparamsBefore) = c("asymptote", "rate constant","AI")        #set column names
rownames(tempOMparamsBefore)[1]=as.character(flyname)              #set row names to flynames
kable(tempOMparamsBefore)                                          #plot table with data

###collect all coefficients in dataframe for the group
if(!exists("OMparamsBefore"))
{
  OMparamsBefore=tempOMparamsBefore
} else {
  OMparamsBefore[as.character(flyname),]=tempOMparamsBefore
}

```

`r if (beforeafter) '### Optomotor parameters at end'`

```{r OMparamsAfter, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    

###generate traces to fit nonlinear model for parameter estimation
# Find the midpoint in left and right turning in averaged traces (OMdata)
OMmidAfter = OMdataAfter$time[nrow(OMdataAfter)]/2

#right turning traces
x1=OMdataAfter$time[OMdataAfter$time<=OMmidAfter]                     #time
y=OMdataAfter[[as.character(flyname)]][OMdataAfter$time<=OMmidAfter]  #mean right OM response
y=rollapply(y, 40,mean, fill="extend")                                #apply rolling average of 20 data points (2s)
xy <-sortedXyData(x1,y)                                               #create sortedXYdata object

LinMod = lm(y ~ x1, data = xy)                                        #fit linear model to OMdata right turning

#plot(xy)
#abline(LinMod)
#LinMod$coefficients[2]

if(abs(LinMod$coefficients[2])<0.0001){                     #if the linear model shows hardly any slope, set asymptote as mean of all values
  right=c(b0=0, b1=mean(y), lrc=-1000)
  } else {
right<-tryCatch(                                            #extract the right turning coefficients into a vector. Set vector to zero for errors
    expr = {
                    right<-NLSstAsymptotic(xy)
                     },
    error = function(e){
                      right=c(b0=0, b1=0, lrc=-1000)        #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(right)
                     },
    warning = function(w){
                      right=c(b0=0, b1=0, lrc=-1000)        #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(right)
                     },
    finally = {}
)}

#left turning traces
x1=OMdataAfter$time[OMdataAfter$time>OMmidAfter]                     #time
x1=x1-x1[1]                                                              #set time to start from zero
y=OMdataAfter[[as.character(flyname)]][OMdataAfter$time>OMmidAfter]   #mean left OM response
y=rollapply(y, 40,mean, fill="extend")                                   #apply rolling average of 20 data points (2s)
xy <-sortedXyData(x1,y)                                                  #create sortedXYdata object

LinMod = lm(y ~ x1, data = xy)                              #fit linear model to OMdata left turning

#plot(xy)
#abline(LinMod)
#LinMod$coefficients[2]

if(abs(LinMod$coefficients[2])<0.0001){                     #if the linear model shows hardly any slope, set asymptote as mean of all values
  left=c(b0=0, b1=mean(y), lrc=-1000)
  } else {
left<-tryCatch(                                             #extract the left turning coefficients into a vector. Set vector to zero for errors
    expr = {
                    left<-NLSstAsymptotic(xy)
                     },
    error = function(e){
                      left=c(b0=0, b1=0, lrc=-1000)         #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(left)
                     },
    warning = function(w){
                      left=c(b0=0, b1=0, lrc=-1000)         #set logarithmic rate constant to -1000 so exp(lrc)=0
                   return(left)
                     },
    finally = {}
)}


###calculate averages and asymmetries to three significant digits and put them into a vector
rightasymp = right[1]+right[2]
leftasymp = left[1]+left[2]
#sanity checks
if(leftasymp < maxfly[1]){leftasymp=0}
if(rightasymp > maxfly[2]){rightasymp=0}

if (leftasymp==0 & rightasymp==0){aveOMcoeff=c(0,0,0)
} else {
      aveOMcoeffAfter=c(signif((abs(rightasymp)+abs(leftasymp))/2, 3),                           #mean asymptote
                   signif(((exp(right[3])+exp(left[3]))/2)*1000, 3),                             #mean rate constant [s]
                   signif((abs(rightasymp)-abs(leftasymp))/(abs(rightasymp)+abs(leftasymp)), 3)) #asymmetry index
}

###make vector into a dataframe for easy exporting and plotting
tempOMparamsAfter <- as.data.frame(rbind(as.numeric(aveOMcoeffAfter))) #convert to dataframe
names(tempOMparamsAfter) = c("asymptote", "rate constant","AI")        #set column names
rownames(tempOMparamsAfter)[1]=as.character(flyname)              #set row names to flynames
kable(tempOMparamsAfter)                                          #plot table with data

###collect all coefficients in dataframe for the group
if(!exists("OMparamsAfter"))
{
  OMparamsAfter=tempOMparamsAfter
} else {
  OMparamsAfter[as.character(flyname),]=tempOMparamsAfter
}

```


# Power spectrum and spectrogram

```{r spectra, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8}
    meanspec(rawdata$fly, f = down_sample_rate, wl = 600) #plot mean power spectrum for each fly
    title(paste(FlyBehavior, "Powerspectrum", flyname))
    spectro(rawdata$fly, f = down_sample_rate, wl=600) #plot spectrogram for each fly
    title(paste(FlyBehavior, "Spectrogram", flyname))
    speclist[[l]] = as.data.frame(meanspec(rawdata$fly, f = down_sample_rate, wl = 600, plot = FALSE)) #collect each spectrum for later averaging
```


```{r PIs, eval=(!all(is.na(sequence$lambda))), echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}
cat("\n# Learning and avoidance scores\n")

      barplot(as.vector(na.omit(sequence$lambda)), main = paste("Performance Indices", flyname),
            xlab="Time [periods]",
            ylab="PI [rel. units]",
            space=0,
            col = as.vector(na.omit(sequence$color)))
```

