---
title: "Single Fly Descriptive Data Evaluation Sheet"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
---
<style type="text/css">
  body .main-container {
    max-width: 1800px !important;
  }
  h1.title {
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<center><h2>for time series experiment `r paste(flyname)`</h2></center>
<center>Rendered on `r format(Sys.time(), '%d %B %Y')`</center>

# Metadata

**Date and time of the experiment:** `r format(as.POSIXct(gsub("T", " ", experiment$dateTime)),'%A, %B %d, %Y %H:%M:%S')`  
**Raw data file (XML):** `r paste('<a href="', basename(xml_name), '">', flyname, '</a>', sep = "")`  
**Experimenter:** `r paste(experimenter$firstname, experimenter$lastname, sep = " ")`; **ORCID:** [`r paste(experimenter$orcid)`](http://orcid.org/`r paste(experimenter$orcid)`)  
**Experiment description:** `r paste(experiment$description)`  
**Experiment duration:** `r paste(experiment$duration)`s  
**Arena type:** `r paste(experiment$arena_type)`; **Measurement device:** `r paste(experiment$meter_type)`  
**Fly description:** `r paste(fly$description)`; **FlyBase ID:** `r hyperlinks.FBids(fly$flybase)`  
**Sampling rate:** `r paste(experiment$sample_rate)`Hz
```{r sr_condition, echo=FALSE}
sampleratecondition <- round(real_sample_rate)>20
```

```{r samplerate, eval=sampleratecondition, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, comment=NA, results = 'asis'}
cat("<b>Actual samplingrate: </b>", paste(signif(real_sample_rate,4)), " Hz; <b>Downsampled sampling rate (target 20 Hz):</b> ", paste(signif(down_sample_rate,4)), " Hz")
```

# Experimental Design
```{r expdesign, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 3, fig.width = (NofPeriods+1)}
    #plot table of period sequence
    periods=t(sequence)
    colnames(periods)<-sprintf("Period %s",1:NofPeriods)
reactable(setDT(as.data.frame(periods), keep.rownames = "property")[],
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  resizable = TRUE,
  columns = list(
  property = colDef(
    style = list(position = "sticky", left = 0, background = "#fff", zIndex = 1),
    headerStyle = list(position = "sticky", left = 0, background = "#fff", zIndex = 1)
  )))    
```

# Time Traces
## Interactive Time Traces

```{r dytracesraw, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 4, fig.width = 12}    
    
    ##dyplot traces
    dytraces(traces)

```

## Interactive Time Traces (Downsampled data)

```{r dytracesdown, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 4, fig.width = 12}    
    
    ##dyplot traces
    dytraces(rawdata)

```

## Time Traces by Period (20Hz)

```{r PeriodTraces, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 10}

#add columns for PIs and categories to sequence data
sequence$lambda <- NA
sequence$category <- NA
  training = NA
for(i in 1:NofPeriods){
  
  #save colors for later plotting
  if(sequence$outcome[i]==0){
    sequence$color[i]="lightyellow"
    sequence$histocolor[i]="darkgreen"
  } else {
    sequence$color[i]="orange"
    sequence$histocolor[i]="orange"
  }
  if(sequence$outcome[i]==0){
    sequence$color[i]="lightyellow"
    sequence$histocolor[i]="darkgreen"
  } else {
    sequence$color[i]="orange"
    sequence$histocolor[i]="orange"
  }
  if(sequence$contingency[i]==""){
    sequence$color[i]=NA
    sequence$histocolor[i]="red"
    sequence$category[i]=NA
 }
  if(sequence$contingency[i]!=""){
    sequence$category[i] <- paste(sequence$pattern[i],sequence$contingency[i])
  } 
#add training/test to the graphs

  if(sequence$outcome[i]==0){
    training[i] =  "Test"
  } else {
    training[i] =  "Training"
  }
  ###divide raw data into period chunks and prepare period data for later use in group evaluations
      temp  <- rawdata[rawdata$period == i, ]
      keeps = c("a_pos","fly")
      
      if((sequence$type[i]=="yt"||sequence$type[i]=="sw") & sequence$contingency[i]=="right_torque")   #inverse torque data for group histograms
      {
        temp_inv = temp
        temp_inv$fly = -temp_inv$fly # invert the torque values
        period.data[[i]] <- temp_inv[keeps] #list only arena position and fly behavior data by period
      } else if((sequence$type[i]=="fs"||sequence$type[i]=="color") & sequence$contingency[i]=="2_4_Q") #shift position data for group histograms
      {
        temp_shifted = temp
        temp_shifted$a_pos = temp_shifted$a_pos + 900 #shift the position values by 90°
        temp_shifted$a_pos[temp_shifted$a_pos>1796] = temp_shifted$a_pos[temp_shifted$a_pos>1796]-3600 #wrap the shifted 90° back around to -90°..-180°
        period.data[[i]] <- temp_shifted[keeps] #list only arena position and fly behavior data by period
      } else period.data[[i]] <- temp[keeps] #keep only arena position and fly behavior data by period
      
      if(any(c("yt","color","sw","fs","optomotorR","optomotorL") %in% sequence$type[i]))
      {
        fly_pos_traces(temp) #plot the fly behavior and arena position time traces
      }

      ##calculate PIs and 
      if(!exists("PIprofile")){PIprofile <- data.frame(matrix(ncol = NofPeriods))} ##create data.frame for adding PIs if it doesn't exist, yet
      if(sequence$type[i]=="fs"||sequence$type[i]=="color") #for arena position
      {
        t1 = sum(abs(temp$a_pos) >= 450 & abs(temp$a_pos) <= 1350)
        t2 = nrow(temp)-t1
        sequence$lambda[i] = (t1-t2)/(t1+t2)
        if (sequence$contingency[i] == '2_4_Q'){sequence$lambda[i]=-sequence$lambda[i]}
        if(l==1){PIprofile[1,i]=sequence$lambda[i]}
      } else if (sequence$type[i]=="yt" || sequence$type[i]=="sw") #calculate torque PIs
      {
        t2 = sum(temp$fly < 0)
        t1 = nrow(temp)-t2
        sequence$lambda[i] = (t1-t2)/(t1+t2)
        if (sequence$contingency[i] == 'right_torque'){sequence$lambda[i]=-sequence$lambda[i]}
        if(l==1){PIprofile[1,i]=sequence$lambda[i]}
      }
      
      ##add categories to separate dataframe
      if(!exists("Categories")){Categories <- data.frame(matrix(ncol = NofPeriods))} #create it, if it doesn't exist, yet and...
      if(l==1){Categories[1,i]=sequence$category[i]} #...add category variable to dataframe
      
      ##add column to rawdata that indicates if the fly was punished or not, leaving out all periods without specified contingency
      if(sequence$contingency[i] == 'left_torque'){rawdata$punished[rawdata$period==i] <- ifelse(rawdata$fly[rawdata$period==i] < 0, 1, 0) 
      }else if(sequence$contingency[i] == 'right_torque') {rawdata$punished[rawdata$period==i] <- ifelse(rawdata$fly[rawdata$period==i] >= 0, 1, 0) 
      }else if(sequence$contingency[i] == '2_4_Q'){rawdata$punished[rawdata$period==i] <- ifelse(abs(rawdata$a_pos[rawdata$period==i]) >= 450 & abs(rawdata$a_pos[rawdata$period==i]) <= 1350, 1, 0)
      }else if(sequence$contingency[i] == '1_3_Q'){rawdata$punished[rawdata$period==i] <- ifelse(abs(rawdata$a_pos[rawdata$period==i]) < 450 | abs(rawdata$a_pos[rawdata$period==i]) > 1350, 1, 0)}
  } #for Number of Periods
```

# Histograms
## Fly Behavior Histograms

```{r FlyHistogram, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 16, fig.width = 8}
  subset = rawdata[c("fly","period","a_pos")]
  subset$period = as.factor(subset$period)
par(mfrow=c(round(NofPeriods/5)*2, 3)) # sets the dimension of the multiplot, row x column
old.par = par(mar = c(4, 4, 1, 0)) # sets the margins for the plots (bottom, left, top and right margins respectively)
for (i in 1:NofPeriods){
  df = subset(subset, subset$period == i) # subset the data for only one period
  hist(x = df$fly, breaks = 100,xlim = maxfly, xlab="", ylab = "", border= sequence$histocolor[i],  col = sequence$histocolor[i], main=paste("Period ", i)) #plots the data as a histogram
#add shading in case of torque/switch learning
    if(sequence$contingency[i]=="left_torque"){          #shade left torque area
    rect(-2000, 0, 0, 500, col=("lightgrey"))
    hist(x = df$fly, breaks = 100,xlim = maxfly, xlab="", ylab = "", border= sequence$histocolor[i],  col = sequence$histocolor[i], main=paste("Period ", i), add = TRUE) #plot new hist
  } else if(sequence$contingency[i]=="right_torque"){  #shade right torque area
    rect(0, 0, 2000, 500, col=("lightgrey"))
    hist(x = df$fly, breaks = 100,xlim = maxfly, xlab="", ylab = "", border= sequence$histocolor[i],  col = sequence$histocolor[i], main=paste("Period ", i), add = TRUE) #plot new hist
  }
  abline(v = 0, col="black", lwd=2, lty=2) # adds a line at 0
  mtext(text = paste(FlyBehavior, "[arb units]"),
        cex = 1,
      side = 1,#side 1 = bottom
      line = 2.5)
  mtext(text = "Frequency",
        cex = 1,
      side = 2, #side 2 = left
      line = 2)
  }
hist(subset$fly, breaks = 100, border = "grey40", col = "grey40", main=paste(flyname, "Total"), xlab="", ylab = "") #histogram of pooled data
  mtext(text = paste(FlyBehavior, "[arb units]"),
        cex = 1,
      side = 1,#side 1 = bottom
      line = 2.5)
  mtext(text = "Frequency",
        cex = 1,
      side = 2, #side 2 = left
      line = 2)

```


```{r PositionHistogram, eval=('fs' %in% sequence$type || 'color' %in% sequence$type), echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}    
      cat("\n## Arena Position Histograms\n")

  subset = rawdata[c("fly","period","a_pos")]
  subset$period = as.factor(subset$period)
par(mfrow=c(round(NofPeriods/2), 2)) # sets the dimension of the multiplot, row x column
old.par = par(mar = c(4, 4, 1, 0)) # sets the margins for the plots (bottom, left, top and right margins respectively)
for (i in 1:NofPeriods){
  df = subset(subset, subset$period == i) # subset the data for only one period
  hist(x = df$a_pos, breaks = 180,xlim = c(-1800,1800), xlab="", ylab = "", border= sequence$histocolor[i],  col = sequence$histocolor[i], main=paste("Period ", i), xaxt='n', yaxs="i") #plots the data as a histogram
#add shading in case of torque/switch learning
  if(sequence$contingency[i]=="1_3_Q"){  #shade first and third quadrant
     rect(-1800, 0, -1350, 800, col=("lightgrey"))
     rect(-450, 0, 450, 800, col=("lightgrey"))
     rect(1350, 0, 1800, 800, col=("lightgrey"))
     hist(x = df$a_pos, breaks = 900,xlim = c(-1800,1800), xlab="", ylab = "", border= sequence$histocolor[i],  col = sequence$histocolor[i], main=paste("Period ", i), add = TRUE) #plots histogram again
  } else {
     rect(-1350,0,-450,800, col=("lightgrey"))
     rect(450,0,1350,800, col=("lightgrey"))
     hist(x = df$a_pos, breaks = 900,xlim = c(-1800,1800), xlab="", ylab = "", border= sequence$histocolor[i],  col = sequence$histocolor[i], main=paste("Period ", i), add = TRUE) #plots histogram again
  }
  abline(v = 0, col="black", lwd=1, lty="dotted") # adds a line at 0
  abline(v = 900, col="black", lwd=1, lty="dotted") # adds a line at 90°  
  abline(v = -900, col="black", lwd=1, lty="dotted") # adds a line at -90° 
  axis(side=1, at=c(-1800, -900, 0, 900, 1800))
  mtext(text = paste("Arena Position [degrees*10]"),
        cex = 1,
      side = 1,#side 1 = bottom
      line = 2.5)
  mtext(text = "Frequency",
        cex = 1,
      side = 2, #side 2 = left
      line = 2)
  }
hist(subset$a_pos, breaks = 360, border = "grey40", col = "grey40", main=paste(flyname, "Total"), xlab="", ylab = "", xaxt='n', yaxs="i") #histogram of pooled data
  axis(side=1, at=c(-1800, -900, 0, 900, 1800))
  #add quadrant borders
  abline(v = -1350, col="black", lwd=1, lty="dotted") # adds a line at -135°
  abline(v = -450, col="black", lwd=1, lty="dotted") # adds a line at -45° 
  abline(v = 450, col="black", lwd=1, lty="dotted") # adds a line at 45° 
  abline(v = 1350, col="black", lwd=1, lty="dotted") # adds a line at 135° 
  
  mtext(text = paste("Arena Position [degrees*10]"),
        cex = 1,
      side = 1,#side 1 = bottom
      line = 2.5)
  mtext(text = "Frequency",
        cex = 1,
      side = 2, #side 2 = left
      line = 2)      

```


<!-- Evaluate optomotor data if they are present -->

```{r om, echo=FALSE}
om <- any(grepl("optomotor", sequence$type)==TRUE)    ###determine if there are optomotor periods
if (om & any(!grepl("optomotor", sequence$type)==TRUE)){   ###if there are non-optomotor periods...
  if (grepl("optomotor", sequence$type[1]) & grepl("optomotor", tail(sequence$type, 1))){ ###...and the optomotor periods are in the beginning and the end
    beforeafter = TRUE
    om = FALSE
  } else {beforeafter = FALSE}
}  else {beforeafter = FALSE}
```

`r if (om | beforeafter) '# Optomotor evaluations'`
`r if (om | beforeafter) '## Optomotor traces (right/left)'`

```{r OptomotorTraces, eval=om, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}    

####### Do not separate optomotor traces at all ##########

### extract optomotor traces and generate dataframe for right (clockwise) turning arena
OMperiods = which(grepl("optomotorR", sequence$type))                                     #collect the right turning OM periods
rightOMdata <- generateOMdata(OMperiods)                                                  #call function which assembles optomotor data for the identified periods
omtitle = "Right (clockwise) arena rotations -"                                           #set plot title
plotrawOMtraces(rightOMdata, omtitle)                                                     #call function to pot the OM traces

### extract optomotor traces and generate dataframe for left turning (counter-clockwise) arena
OMperiods = which(grepl("optomotorL", sequence$type))                                     #collect the left turning OM periods
leftOMdata <- generateOMdata(OMperiods)                                                   #call function which assembles optomotor data for the identified periods
leftOMdata$time=leftOMdata$time+max(rightOMdata$time)                                     #make left data start after right data
omtitle = "Left (counter-clockwise) arena rotations -"                                    #set plot title
plotrawOMtraces(leftOMdata, omtitle)                                                      #call function to pot the OM traces

### generate averaged optomotor responses over both rotation directions
tempOMdata=rbind(rightOMdata,leftOMdata)                                                  #combine left and right OMdata in one dataframe
tempOMdata=aggregate(tempOMdata$fly,FUN=mean,by=list(tempOMdata$time))                    #average across periods
colnames(tempOMdata) <- c("time", as.character(flyname))                                  #set dataframe names

if(!exists("OMdata"))
{
  OMdata=tempOMdata
} else {
  OMdata[[as.character(flyname)]]=tempOMdata[[as.character(flyname)]]
}
```

`r if (beforeafter) '### Optomotor traces at start'`

```{r OptomotorTracesBefore, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}    

######### Separate optomotor evaluations: optomotor raw traces at the start of the experiment ###############

### extract and plot optomotor traces for right (clockwise) turning arena at start of experiment
OMperiods = which(grepl("optomotorR", head(sequence$type)))                               #collect the right turning OM periods in the first half of the experiment
rightOMdataBefore <- generateOMdata(OMperiods)                                            #call function which assembles optomotor data for the identified periods
omtitle = "Right (clockwise) arena rotations -"                                           #set plot title
plotrawOMtraces(rightOMdataBefore, omtitle)                                               #call function to pot the OM traces

### extract optomotor traces and generate dataframe for left turning (counter-clockwise) arena at start of experiment
OMperiods = which(grepl("optomotorL", head(sequence$type)))                               #collect the left turning OM periods in the first half of the experiment
leftOMdataBefore <- generateOMdata(OMperiods)                                             #call function which assembles optomotor data for the identified periods
leftOMdataBefore$time=leftOMdataBefore$time+max(rightOMdataBefore$time)                   #make left data start after right data
omtitle = "Left (counter-clockwise) arena rotations -"                                    #set plot title
plotrawOMtraces(leftOMdataBefore, omtitle)                                                #call function to pot the OM traces

### generate averaged optomotor responses over both rotation directions
tempOMdataBefore=rbind(rightOMdataBefore,leftOMdataBefore)                                #combine left and right OMdata in one dataframe
tempOMdataBefore=aggregate(tempOMdataBefore$fly,FUN=mean,by=list(tempOMdataBefore$time))  #average across periods
colnames(tempOMdataBefore) <- c("time", as.character(flyname))                            #set dataframe names

if(!exists("OMdataBefore"))
{
  OMdataBefore=tempOMdataBefore
} else {
  OMdataBefore[[as.character(flyname)]]=tempOMdataBefore[[as.character(flyname)]]
}
```

`r if (beforeafter) '### Optomotor traces at end'`

```{r OptomotorTracesAfter, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}    

######### Separate optomotor evaluations: optomotor raw traces at the end of the experiment ###############

### extract optomotor traces and generate dataframe for right (clockwise) turning arena at the end of the experiment
OMperiods = which(grepl("optomotorR", tail(sequence$type, 8)))+NofPeriods-8               #collect the right turning OM periods in the last half of the experiment
rightOMdataAfter <- generateOMdata(OMperiods)                                             #call function which assembles optomotor data for the identified periods
omtitle = "Right (clockwise) arena rotations -"                                           #set plot title
plotrawOMtraces(rightOMdataAfter, omtitle)                                                #call function to pot the OM traces

### extract optomotor traces and generate dataframe for left turning (counter-clockwise) arena at end of experiment
OMperiods = which(grepl("optomotorL", tail(sequence$type, 8)))+NofPeriods-8               #collect the left turning OM periods in the last half of the experiment
leftOMdataAfter <- generateOMdata(OMperiods)                                              #call function which assembles optomotor data for the identified periods
leftOMdataAfter$time=leftOMdataAfter$time+max(rightOMdataAfter$time)                      #make left data start after right data
omtitle = "Left (counter-clockwise) arena rotations -"                                    #set plot title
plotrawOMtraces(leftOMdataAfter, omtitle)                                                 #call function to pot the OM traces

### generate averaged optomotor responses over both rotation directions
tempOMdataAfter=rbind(rightOMdataAfter,leftOMdataAfter)                                   #combine left and right OMdata in one dataframe
tempOMdataAfter=aggregate(tempOMdataAfter$fly,FUN=mean,by=list(tempOMdataAfter$time))     #average across periods
colnames(tempOMdataAfter) <- c("time", as.character(flyname))                             #set dataframe names

if(!exists("OMdataAfter"))
{
  OMdataAfter=tempOMdataAfter
} else {
  OMdataAfter[[as.character(flyname)]]=tempOMdataAfter[[as.character(flyname)]]
}
```

`r if (om | beforeafter) '## Averaged optomotor response'`

```{r OptomotorAverage, eval=om, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}    

plotaveOMtraces(OMdata)
```

`r if (beforeafter) '### Averaged optomotor response at start'`

```{r OptomotorAverageBefore, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}    

plotaveOMtraces(OMdataBefore)
```

`r if (beforeafter) '### Averaged optomotor response at end'`

```{r OptomotorAverageAfter, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}    

plotaveOMtraces(OMdataAfter)
```

`r if (om | beforeafter) '## Further optomotor evaluations'`

```{r moreOptomotor, eval=om, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 8, fig.width = 6, comment=NA, results='asis'}    

####### Do not separate optomotor parameters at all ##########

tempOMparams <- OMparamextract_new(OMdata) #call fucntion for parameter estimation
kable(tempOMparams)  #plot table with data

###collect all coefficients in dataframe for the group
if(!exists("OMparams"))
{
  OMparams=tempOMparams
} else {
  OMparams[as.character(flyname),]=tempOMparams
}

```

`r if (beforeafter) '### Optomotor parameters at start'`

```{r OMparamsBefore, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}    

######### Separate optomotor evaluations: optomotor parameters at the start of the experiment ###############

tempOMparamsBefore<-OMparamextract(OMdataBefore) #call fucntion for parameter estimation
kable(tempOMparamsBefore)  #plot table with data

###collect all coefficients in dataframe for the group
if(!exists("OMparamsBefore"))
{
  OMparamsBefore=tempOMparamsBefore
} else {
  OMparamsBefore[as.character(flyname),]=tempOMparamsBefore
}

```

`r if (beforeafter) '### Optomotor parameters at end'`

```{r OMparamsAfter, eval=beforeafter, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}    

######### Separate optomotor evaluations: optomotor parameters at the end of the experiment ###############

tempOMparamsAfter<-OMparamextract(OMdataAfter) #call fucntion for parameter estimation
kable(tempOMparamsAfter)  #plot table with data

###collect all coefficients in dataframe for the group
if(!exists("OMparamsAfter"))
{
  OMparamsAfter=tempOMparamsAfter
} else {
  OMparamsAfter[as.character(flyname),]=tempOMparamsAfter
}

```


# Power spectrum and spectrogram

```{r spectra, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8}
    meanspec(rawdata$fly, f = down_sample_rate, wl = 600) #plot mean power spectrum for each fly
    title(paste(FlyBehavior, "Powerspectrum", flyname))
    spectro(rawdata$fly, f = down_sample_rate, wl=600) #plot spectrogram for each fly
    title(paste(FlyBehavior, "Spectrogram", flyname))
    speclist[[l]] = as.data.frame(meanspec(rawdata$fly, f = down_sample_rate, wl = 600, plot = FALSE)) #collect each spectrum for later averaging
```


```{r dwellingtimes, eval=Dwell, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 12, fig.width = 8, comment=NA, results='asis'}

cat("\n# Dwelling Times\n")

#split the punishment data into periods
dwellperiod <- split(rawdata$punished, rawdata$period) #split the punished/unpunished datapoints by period
c=0 #counter to arrange figures

for (i in nonOMperiods) {
    c=c+1 #increment counter
    dwellperiod[[i]] = rle(dwellperiod[[i]])            #run length encode the punished and unpunished data points
    dwellperiod[[i]][[1]] = dwellperiod[[i]][[1]]*0.05  #convert run lenghts into seconds. Now we have the raw dwelling times.
    
    if(is.null(unlist(split(dwellperiod[[i]][[1]],dwellperiod[[i]][[2]])["0"]))){dwelldata$unpunished[[i]]=as.integer(0) #if no data points, set to zero seconds
    }else{dwelldata$unpunished[[i]] = unlist(split(dwellperiod[[i]][[1]],dwellperiod[[i]][[2]])["0"])} #unpunished dwelling times
      
    if(is.null(unlist(split(dwellperiod[[i]][[1]],dwellperiod[[i]][[2]])["1"]))){dwelldata$punished[[i]]=as.integer(0) #if no data points, set to zero seconds
    } else{dwelldata$punished[[i]] = unlist(split(dwellperiod[[i]][[1]],dwellperiod[[i]][[2]])["1"])}  #punished dwelling times

#calculate mean dwelling times and add them to the list used in project evaluations    
    dwellmeans$unpunished[l,i] =  mean(dwelldata$unpunished[[i]])
    dwellmeans$punished[l,i] =  mean(dwelldata$punished[[i]])

#Prepare data for plotting cumulative percentages
#sort dwelling times for x-axis for cumulative distributions    
    DwellTpunished = sort(unique(dwelldata$punished[[i]]))
    DwellTunpunished = sort(unique(dwelldata$unpunished[[i]]))
#calculate the cumulative percentages and shift the data down. Shift moves the data down by the number specified in n, fill=100 replaces the shifted value with 100, type="lag" decides in which direction to shift the data.
    dwelldata$unpunished[[i]] = melt(shift((100-freq(dwelldata$unpunished[[i]], cum = TRUE)$`val%cum`), n=1, fill=100, type="lag")) 
    dwelldata$punished[[i]] = melt(shift((100-freq(dwelldata$punished[[i]], cum = TRUE)$`val%cum`), n=1, fill=100, type="lag"))
#move x-axes to shifted/melted dwelldata dataframe
    dwelldata$punished[[i]]$times = DwellTpunished
    dwelldata$unpunished[[i]]$times = DwellTunpunished
  
#plots the cumulative percentage data
  dwellplots[[c]] =ggplot() +
    geom_line(data= dwelldata$unpunished[[i]], aes(x=times, y=value,color="red"))+
    geom_point(data= dwelldata$unpunished[[i]], aes(x=times, y=value,color="red"))+
    geom_line(data= dwelldata$punished[[i]], aes(x=times, y=value,color="blue"))+
    geom_point(data= dwelldata$punished[[i]], aes(x=times, y=value,color="blue"))+
    scale_y_log10( breaks=c(0,10,100))+
    scale_x_log10() +
    ggtitle(paste(flyname, "Period", i))+
    theme_light() +
    labs(x="time [sec]", y="% events",  color = "Outcome") +
    annotate("text", Inf, Inf, label = training[i], hjust =1.05, vjust = 1)+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
    labs(color='Outcome') +
    scale_color_manual(values=c("firebrick3", "steelblue4"),
                       name="Experimental\nCondition",
                       labels=c("Punished", "Unpunished"))+
    theme(plot.title = element_text(color="black", size=12, face="bold", hjust = 0.5))
}
grid.arrange(grobs = dwellplots, ncol=2)

```

```{r PIs, eval=(!all(is.na(sequence$lambda))), echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.align = "center", fig.height = 6, fig.width = 8, comment=NA, results='asis'}
cat("\n# Learning and avoidance scores\n")

      barplot(as.vector(na.omit(sequence$lambda)), main = paste("Performance Indices", flyname),
            xlab="Time [periods]",
            ylab="PI [rel. units]",
            space=0,
            col = as.vector(na.omit(sequence$color)))
```
